---
title: "Introduction to interactive documents"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Shiny Articles: [Introduction to interactive](documentshttp://shiny.rstudio.com/articles/interactive-docs.html)

You can embed Shiny components in an R Markdown file to create an interactive report or slide show.

You can make an R markdown document interactive in two steps:

1. add `runtime: shiny` to the document's YAML header.
2. add Shiny widgets and Shiny render functions to the file's R code chunks.

### An interative shiny example:

#### Here are two Shiny widgets

```{r echo = FALSE}
selectInput("n_breaks", label = "Number of bins:",
            choices = c(10, 20, 35, 50), selected = 20)

sliderInput("bw_adjust", label = "Bandwidth adjustment:",
            min = 0.2, max = 2, value = 1, step = 0.2)
```

#### ... that build a histogram

```{r echo = FALSE}
renderPlot({
    hist(faithful$eruptions, probability = TRUE,
         breaks = as.numeric(input$n_breaks),
         xlab = "Duration (minutes)",
         main = "Geyser eruption duration")
    
    dens <- density(faithful$eruptions, adjust = input$bw_adjust)
    lines(dens, col = "blue")
})
```

### The structure of an interactive document

When you run an interactive document, `rmarkdown` extracts the code in your code chunks and places them into a pseudo **server.R** file. R Markdown uses the html output of the markdown file as an **index.html** file to place the reactive elements into.

As a result, outputs in one code chunk an use widgets and reactive expressions that occur in other code chunks.

Since R Markdown document provides a layout for the app, you do not need to write an ui.R file.

### Sharing interactive documents

A number of ways to share:

1. Email .Rmd and run in the Rstudio
2. Host the document with Shiny Server
3. Host the document at ShinyApps.io

**Note**: interactive documents cannot be shared as a standalone HTML file.



